From Freebase to Wikidata: The Great Migration

### 2015-10-18 paper.tex
comments from danbri

One such collaborative knowledge base is Freebase,
publicly launched by Metaweb in 2007 and acquired by Google in 2010.
Another example is Wikidata, a~collaborative knowledge base developed by
Wikimedia Deutschland since 2012 and operated by the Wikimedia Foundation.
Due to the success of Wikidata, Google announced in 2014 to shut down Freebase and
help the community with the transfer of Freebase's content to Wikidata~\cite{freebase2014shutdown}.

@@

Freebase\footnote{Freebase: \url{https://www.freebase.com}} is an open and
collaborative knowledge base publicly launched in 2007 by Metaweb and 
acquired in 2010 by Google~\cite{freebase}.
It was used as the open core of the Google Knowledge Graph,
and has found many use cases outside of Google.
Due to the success of Wikidata,
Google announced in 2014 to close Freebase and offer most of its content to the
Wikidata community~\cite{freebase2014shutdown,freebase-end}.

@@

The second challenge is that the Wikidata community is very eager to have references
for their statements, \emph{i.e.}, sources that Freebase usually did not store,
except for some specific data like populations and unemployment rates.
For these two specific cases, we just map the ``\texttt{source}''
\footnote{The two main properties used in Freebase for sources are
\texttt{/measurement\_unit/dated\_integer/source} and
\texttt{/measurement\_unit/dated\_percentage/source}}
Freebase properties included the CVTs encoding these data into Wikidata references.
\todo{Denny: Could you find a better formulation?}

@@

The data quality of Freebase was examined thoroughly by the Wikidata community
and anecdotal evidence indeed confirmed that there were certain challenges.
For example, a~small city in France was assigned an ISO country code,%
\footnote{Freebase topic page of the city of Saint-Martin: \url{http://www.freebase.com/m/03cc0d0}}
then Boston (the city in Massachusetts) had the type \texttt{/people/person},
and several other occurrences of data quality issues were revealed.
In consequence, a~fully automatic upload of all the content from Freebase into Wikidata
did not seem advisable as the expectations of the Wikidata community regarding the
quality of automatically uploaded data are high (and rightly so).
Despite the fact that an automated upload of all the data
would have led to accelerated growth in the short term,
such a~fast increase of available data would certainly have endangered
the sense of ownership of the Wikidata community for its data.
As an alternative approach, we decided to rely on crowdsourced human curation
and created the \emph{Primary Sources Tool},
a~widget that displays Freebase statements
for curation by the contributor that can be added to the currently shown Wikidata item.
The tool will be described in more detail in \autoref{sec:primary-sources-tool}.

@@

\textbf{Schema mappings.}
There has been extensive research on the topic of mapping the structures
of two databases or knowledge bases
(some surveys are~\cite{mappingsurvey1,mappingsurvey2,mappingsurvey3}).
The current work is, in many technical respects, rather simple:
we are looking for a~one-directional mapping (since the case of data
flowing from Wikidata back to Freebase is not relevant),
a~one-off data flow (since Freebase is now read-only),
mostly individually evaluated by human contributors
(therefore not requiring a perfect mapping).
The effort most similar to ours is the DBpedia Mapping Wiki~\cite{lehmann2014dbpedia}.
The DBpedia community is manually writing and maintaining the mapping from
the DBpedia extraction process (working on Wikipedia) to the DBpedia ontology.
The main difference is that the DBpedia mapping is running continuously
and does not allow for point-fixes, thus necessitating a higher quality of the mapping,
whereas for the Freebase to Wikidata mapping every single results is expected to be curated.

>>>>

One such collaborative knowledge base is Freebase,
publicly launched by Metaweb in 2007 and acquired by Google in 2010.
Another example is Wikidata, a~collaborative knowledge base developed by
Wikimedia Deutschland since 2012 and operated by the Wikimedia Foundation.
Due to the success of Wikidata, Google announced in 2014 their intent to shut down Freebase and
help the community with the transfer of Freebase's content to Wikidata~\cite{freebase2014shutdown}.

@@

Freebase\footnote{Freebase: \url{https://www.freebase.com}} is an open and
collaborative knowledge base publicly launched in 2007 by Metaweb and 
acquired in 2010 by Google~\cite{freebase}.
It was used as the open core of the Google Knowledge Graph,
and has found many use cases outside of Google.
Due to the success of Wikidata,
Google announced in 2014 its intent to close Freebase and help with
the migration of the content to Wikidata~\cite{freebase2014shutdown,freebase-end}.

@@

The second challenge is that the Wikidata community is very eager to have references
for their statements, \emph{i.e.}, sources that Freebase usually did not store,
except for some specific data like populations and unemployment rates.
For these two specific cases, we map the Freebase properties describing sources,%
\footnote{The two main properties used in Freebase for sources are
\texttt{/measurement\_unit/dated\_integer/source} and
\texttt{/measurement\_unit/dated\_percentage/source}}
expressed as CVTs, into Wikidata references on the appropriate claim.

@@

The data quality of Freebase was discussed by the Wikidata community, and it
was decided that the quality overall was not sufficient for a direct import.
For example, a~small city in France was assigned an ISO country code,%
\footnote{Freebase topic page of the city of Saint-Martin: \url{http://www.freebase.com/m/03cc0d0}}
then Boston (the city in Massachusetts) had the type \texttt{/people/person},
and several other occurrences of data quality issues were revealed.
In consequence, a~fully automatic upload of all the content from Freebase into Wikidata
did not seem advisable as the expectations of the Wikidata community regarding the
quality of automatically uploaded data are high (and rightly so).
Despite the fact that an automated upload of all the data
would have led to accelerated growth in the short term,
such a~fast increase of available data would certainly have endangered
the sense of ownership of the Wikidata community for its data.
As an alternative approach, we decided to rely on crowdsourced human curation
and created the \emph{Primary Sources Tool},
a~widget that displays Freebase statements
for curation by the contributor that can be added to the currently shown Wikidata item.
The tool will be described in more detail in \autoref{sec:primary-sources-tool}.

@@

\textbf{Schema mappings.}
There has been extensive research on the topic of mapping the structures
of two databases or knowledge bases
(some surveys are~\cite{mappingsurvey1,mappingsurvey2,mappingsurvey3}).
The current work is, in many technical respects, rather simple:
we are looking for a~one-directional mapping (since the case of data
flowing from Wikidata back to Freebase is not relevant),
a~one-off data flow (since Freebase is now read-only),
mostly individually evaluated by human contributors
(therefore not requiring a perfect mapping).
The effort most similar to ours is the DBpedia Mapping Wiki~\cite{lehmann2014dbpedia}.
The DBpedia community is manually writing and maintaining the mapping from
the DBpedia extraction process (working on Wikipedia) to the DBpedia ontology.
The main difference is that the DBpedia mapping is running continuously
and does not allow for point-fixes, thus necessitating a higher quality of the mapping,
whereas for the Freebase to Wikidata mapping every single results is expected to be curated.
The schema.org structured data project is also exploring mappings to and from Wikidata,
to facilitate the use of Wikidata's vocabulary as schema.org markup across the wider Web.
This involves similar issues mapping schema.org's CVT-like statement qualification mechanism, Roles,
to Wikidata conventions.%
\footnote{On Roles: \url{http://schema.org/Role}. Mappings from Schema.org to Wikidata:
\url{https://github.com/schemaorg/schemaorg/issues/280}. On the interaction of Schema.org
and Wikidata: \url{https://meta.wikimedia.org/wiki/Wikidata/Notes/Schema.org_and_Wikidata}}
